---
title: My dynamic report
output: html_fragment
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, message=FALSE, warning=FALSE, error=FALSE)

```

```{css}

```

## Bus trip waiting time and crowding patterns

This dynamic report summarises observational data collected from bus trips. It focuses on waiting time, crowding level, and when the trip was observed.

```{r}
library(tidyverse)

csv_url <- "https://docs.google.com/spreadsheets/d/e/2PACX-1vSq8sCLbN7Gk1WVCHG2rTKcyvYiCdYiMqBJAQurDNrVJkV_MM361DtHaR3KTDorBGXHnYUZSMvqa48Q/pub?gid=1103664544&single=true&output=csv"

temp_csv <- tempfile(fileext = ".csv")
options(timeout = 120)
download.file(csv_url, temp_csv, mode = "wb", method = "libcurl", quiet = TRUE)

logged_data <- read_csv(temp_csv, show_col_types = FALSE)

latest_data <- logged_data %>%
  rename(
    timestamp = Timestamp,
    time_of_day = `What time of day was this bus trip?`,
    day_type = `Which day type was this observation recorded on?`,
    wait_minutes = `How many minutes did you wait for the bus?`,
    crowding_level = `How crowded was the bus when you boarded?`,
    on_time_status = `Was the bus on time based on the scheduled arrival?`
  ) %>%
  mutate(
    time_of_day = case_when(
      time_of_day == "8:00" ~ "6:00-8:59 am",
      TRUE ~ time_of_day
    ),
    time_of_day = factor(
      time_of_day,
      levels = c(
        "6:00-8:59 am",
        "9:00-11:59 am",
        "12:00-2:59 pm",
        "3:00-5:59 pm",
        "6:00 pm or later"
      )
    ),
    day_type = factor(
      day_type,
      levels = c("Weekday", "Weekend")
    ),
    crowding_level = factor(
      crowding_level,
      levels = c(
        "Mostly empty",
        "Some seats available",
        "Few seats available",
        "Standing room only",
        "Completely full / very crowded"
      )
    ),
    on_time_status = factor(
      on_time_status,
      levels = c("Early", "On time", "Late", "Not sure")
    )
  )

total_obs <- nrow(latest_data)
mean_wait <- mean(latest_data$wait_minutes)
max_wait <- max(latest_data$wait_minutes)

weekday_mean <- latest_data %>%
  filter(day_type == "Weekday") %>%
  summarise(value = mean(wait_minutes)) %>%
  pull(value)

weekend_mean <- latest_data %>%
  filter(day_type == "Weekend") %>%
  summarise(value = mean(wait_minutes)) %>%
  pull(value)

most_common_crowding <- latest_data %>%
  count(crowding_level, sort = TRUE) %>%
  slice(1) %>%
  pull(crowding_level)




```
A total of `r total_obs` bus trip observations were recorded.  
The average waiting time was `r round(mean_wait, 1)` minutes, and the longest waiting time was `r max_wait` minutes.

Weekday trips had an average waiting time of `r round(weekday_mean, 1)` minutes, while weekend trips had an average waiting time of `r round(weekend_mean, 1)` minutes.

The most common crowding level recorded was **`r most_common_crowding`**.
```{r}
ggplot(latest_data, aes(x = time_of_day)) +
  geom_bar() +
  labs(
    title = "Number of bus trips by time of day",
    x = "Time of day",
    y = "Number of observations"
  ) +
  theme_minimal()

```

This bar chart shows how the observations were distributed across different times of day.
```{r}
ggplot(latest_data, aes(x = crowding_level)) +
  geom_bar() +
  labs(
    title = "Number of bus trips by crowding level",
    x = "Crowding level",
    y = "Number of observations"
  ) +
  coord_flip() +
  theme_minimal()

```

This bar chart shows which crowding levels were observed most often.
```{r}
ggplot(latest_data, aes(x = day_type, fill = on_time_status)) +
  geom_bar(position = "dodge") +
  labs(
    title = "Bus punctuality by day type",
    x = "Day type",
    y = "Number of observations",
    fill = "On-time status"
  ) +
  theme_minimal()

```

This third chart compares punctuality categories between weekdays and weekends.
